A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.

### 2014-03-03 response/response.tex

  \subsection*{Smaller issues}
  \begin{enumerate}

  \item
	Pg. 2: In line 14, explain what does the term “algorithmically consistent” mean.

{\color{red}
  \begin{itemize}
     \item

      A brief definition for the convenience of the reader was added in the
      manuscript. But, algorithmically consistent is a fairly common term in
      plasticity/mechanics. This same meaning is used for 'consistent' refering
      to 'tangent matrices' in \cite{perez2012numerical, perez2000numerical}. It means
      consistent with integrating the plasticity model under the Khun-Tucker
      constraints. Integrating the plasticity model means, given a strain
      increment, determining a stress that does not exceed the yield stress,
      and by doing so computing a linear slope, called a consistent tangent
      modulus, that relates the strain increment to a correct stress increment.
      An alogorithmically consistent tangent stiffness matrix is then one that
      uses the consistent tangent modulii, making it a more accurate
      approximation of the stiffness of the system.

@@

  \item
    Pg. 16: In lines 54-56, you say “It should be pointed out that compared to the (cid:96)2 -
    norms of any of the TS matrices by themselves, the magnitude of the accuracy
    metric ... is relatively small.” Please include in the manuscript the explicit informa-
    tion on the relative error. It seems that you already have such information in the
    paper repository.

{\color{red}  
\begin{itemize}
     \item
    The manuscript has been corrected. 
    \end{itemize}}

  \item
    Pg. 17: Include the values of the probe distances h used for each method.

{\color{red}  
\begin{itemize}
     \item
    Correction made in the manuscript, values used are collected and the reference to the material explaining these values was made more clear.
    %
    The probe distances used in Peridigm are automatically determined with a
    heuristic algorithm and are not set by the user. The algorithm for this in
    \emph{Peridigm} is the used for another Sandia software project called
    \emph{Adagio} \cite{ref-Adaggio}. The probe distances are 1.E-6 times the node spacing for FD and CD which is indended to prevent the onset of roundoff error.
    1.E-100 was hard-coded for CS into the CS method implementation, while AD had no probe distance.
  \end{itemize}}

@@

  \item
    Pg. 17: In lines 53-54, you say “Accuracy trends matched those for the single core
    tests.” Indeed, comparing Figs. 3 and 4, the results for FD seem to be quite the
    same in serial and in parallel. For CS the accuracy seems higher in serial. For CD,
    the accuracy is also higher in serial and, in serial, it has a ﬂuctuating behavior. Can
    you comment on these observations?

{\color{red}  
\begin{itemize}
     \item
     	Manuscript has been updatded. When accuracy is looked at in terms of relative error, where relative error is the accuracy metric previously plotted in the figures divided by the Frobenius norm of the AD TS matrix alone times 100, there is actually little fluctuation, between tests or between serial and parallel tests. The order of magnitude for this measure remains steady for all the methods.
  \end{itemize}}

@@

  \item
    Pgs. 17-18: What are the units of K ? If F int in Eq. (4) etc. is internal force, then K
    should have units of force/length. However, in Figs. 3 and 4 you use the unit of
    MPa, which is force/length2 . Please clarify.

{\color{red}  
\begin{itemize}
     \item
    The manuscript has been corrected. The units of K are actually force / length4, however there is no intuitive relationship between the norm of the TS matrix and some physical characteristic of the system being modeled.
  \end{itemize}}

@@

 \item
    Pg. 24: In line 23, you may add the truncation error O(h2 ) to the expression. Oth-
    erwise, you may need to replace the “=” sign by the “≈” sign. However, adding
    the error term would make more clear the limiting expression when h → 0. Also,
    replace the comma by a period.

{\color{red}  
\begin{itemize}
     \item
         Manuscript corrected.
  \end{itemize}}

\item    
Pg. 24: Why do we need S and X in Appendix B?

     All the members of set X need to be members of set S otherwise the
     approximations don't work. X are numbers from the 'real' world and S are
     numbers from the computer world. For example:

     \begin{verbatim}
     "http://en.wikipedia.org/wiki/IEEE_floating_point" 
     \end{verbatim}

     shows that computers
     and languages following IEEE standards don't represent very large numbers
     with very fine precision at the same time due to memory constraints, and
     there are upper and lower bounds on the magnitudes of nonzero numbers that
     can be represented. Obviously, approximation error is expected, but the
     fact that a user doesn't actually get 'all' the numbers is something that
     is not obvious.

@@

     shows that computers
     and languages following IEEE standards don't represent very large numbers
     with very fine precision at the same time due to memory constraints, and
     there are upper and lower bounds on the magnitudes of nonzero numbers that
     can be represented. Obviously, approximation error is expected, but the
     fact that a user doesn't actually get 'all' the numbers is something that
     is not obvious.

  \end{itemize}}

\item
    Pgs. 25-26: Are the functions in the second column of the table in 2(c) (page 25)
    approximations? Why do you label the column “Approximates Function”? Simi-
    larly, in line 16 (page 26) you say “approximated by the partial derivative”; is it an
    approximation?

      Strictly speaking, a computer can only provide finite precision
      approximations, of approxmations, of functions. Approximations of
      derivatives of functions are also approximations, therefor the label is
      correct in both cases. For example:

      \begin{verbatim}
      "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S"
      \end{verbatim}

      is the implementation of the $exp(x)$ function on 80-bit extended
      precision floating point numbers in the popular GCC compiler. This
      implementation is the most accurate that GCC provides for the $e$
      function on real numbers, but ultimately is still a discrete numerical
      algorithm.

@@

      is the implementation of the $exp(x)$ function on 80-bit extended
      precision floating point numbers in the popular GCC compiler. This
      implementation is the most accurate that GCC provides for the $e$
      function on real numbers, but ultimately is still a discrete numerical
      algorithm.

  \end{itemize}}

  \item
Pg. 26: Please clarify whether storage is needed in automatic differentiation. For
example, do we need to store w, v , and u?

{\color{red}  
\begin{itemize}
     \item
      Clarification is added in the manuscript.  
  \end{itemize}}

  \end{enumerate}

\bibliographystyle{elsarticle-num}
\bibliography{all}
\end{document}

### 2014-03-03 jacobian_calc_CMAME.tex
completed first draft of small issues section

The methods were compared by running two sets of test problems where each
method would solve a problem concurrently. In this context, concurrently means
successively, drawing from the same independent variables, yet within the same
computational process; this definition is characterized in further detail in
Section~\ref{JGAM}. Each test problem simulated a $\SI{4}{\meter}$ by
$\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing
tension along the axis parallel to the long dimension of the block. The test
problems were set up in \emph{Perdigm} as quasistatic equilibrium problems,
where the displacement boundary conditions used to apply tension were applied
gradually in ``load steps'', and an equilibrium solution for each load step was
achieved before applying the next load step. The purpose of applying the load
in gradual steps is a widely used technique to ensure the Newton's method is
always initialized near the actual solution and therefore likely to converge.
The fictional material represented in the model had values of
$\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$
for shear modulus. The peridynamic horizon, a length scale that effectively
sets the bandwidth of tangent-stiffness in the computation, used in the test
problems was always set to three times the nominal discretization size (i.e.
node spacing in the particle discretization scheme). This results in a minimum
tangent-stiffness matrix bandwidth of 21 for a 3D problem (for an optimal node
numbering scheme). Pictured in Fig.~\ref{fig:SCMesh} is a viewgraph for one of
the peridynamic meshes used in this study. The coloration in the image
indicates node numbering, where a smooth color transition means that the nodes
have close ID numbers. The abrupt color changes in the center and at the ends
of the body show that numbering is not contiguous and is assigned with respect to four regions,
called \emph{material blocks} in \emph{Peridigm}. The purpose of designating
these material blocks, for this study, is to register the blocks on the ends of
the mesh as volumetric regions over which to apply displacement boundary
conditions called \emph{volume constraints} in peridynamics.

@@

\begin{table*}[!ht]    
  \scriptsize
  \centering
        \caption{Test Parameters} \label{tab:TestParams}   
       \begin{tabular}{c c c c c c c}
         \toprule
         Peri. nodes & Nozero TS elems. & Nonzero Fraction & Horizon & Load rate & # Load steps & Abs. tolerance \\ 
                     & $\times 10^6$    &                  &\si{\meter}& \si{\centimeter\per\second} &   &     \\
        \midrule
        1000 & $\num{1.99}$             &                   & \num{.3}   & 2.0 & 2 & \num{10^-8}\\
        2000 & $\num{4.25}$             &                   &            &     &   &            \\
        3000 & $\num{7.79}$             &                   &            &     &   &            \\
        4000 & $\num{15.4}$             &                   &            &     &   &            \\
        6000 & $\num{20.0}$             &                   &            &     &   &            \\
        8000 & $\num{31.4}$             &                   &            &     &   &            \\
        12000 & $\num{40.1}$            &                   &            &     &   &            \\
        16000 & $\num{83.9}$            &                   &            &     &   &            \\
        32000 & $\num{165.0}$           &                   &            &     &   &            \\
        1000000 & $\num{1670}$          &                   & \num{.03}  & 2.0 & 2 & \num{10^-8}\\
        \bottomrule
    \end{tabular}
\end{table*}

@@

Data for accuracy measurements were taken according to the methods mentioned in
Section~\ref{tcs}, and reduced using Eqn.~(\ref{eqn:accuracy}). The
averaged results of these accuracy measurements for the serial tests can be
seen in Fig.~\ref{fig:serial_accuracy}. Averaged Frobenius norm (referred to as
$l^2$-norm in the figure labels for conciseness) of the element-wise difference
between a given TS matrix and the TS matrix produced by the AD method is
plotted as a function of the number of nonzero TS matrix elements. CS was shown
to be the most accurate of the methods when compared to AD and was roughly
\emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of
magnitude more accurate than FD}. This accuracy ranking order was maintained
for each test in the series. It should be pointed out that compared to the
$l^2$-norms of any of the TS matrices by themselves, the magnitude of the
accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods
is relatively small. These values for the $l^2$-norms of the TS matrices appear
in the simulation data which is available in the paper repository. The effect
of these accuracy differences on convergence rate of the Newton solver was studied and the
results are reported in section~\ref{sec:PeridigmConvergenceStudy}.  
%
The accuracy of FD and CD could be improved by using a smaller probe distance,
$h$; however, if $h$ is too small, then the methods can suffer from severe
effects, including solution divergence, of the round-off error associated with
subtracting two numbers that are very close to one another.  The value of $h$
used in this study is the default value used in \emph{Peridigm} which is
heuristically derived based on computational node spacing.  One of the beauties
of the CS method is that since dependence on $h$ decreases for small $h$, which
a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$,
cannot be too small, therefore it can be set to machine epsilon and the
analysis performed without the issues associated with round-off error. The
value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to
demonstrate this.
%
\begin{figure}[tbp] \centering
\scalebox{1.0}{\input{./figs/serial_accuracy.pgf}} \caption{Serial test series
accuracy measurements.} \label{fig:serial_accuracy} \end{figure}
%
The averaged results of the accuracy measurements taken for the parallel test
series can be seen in Fig.~\ref{fig:multi_accuracy}. The ordinate axis units
remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the abscissa now
indicates number of compute cores used to solve the test problem. Accuracy
trends matched those for the single core tests. 
%
\begin{figure}[tbp] \centering
\scalebox{1.0}{\input{./figs/multi_accuracy.pgf}} \caption{Multicore test
series accuracy measurements.} \label{fig:multi_accuracy} \end{figure}

@@

@@

\appendix
\renewcommand*{\thesection}{\Alph{section}}
%
\section{Complex-Step Example}
\label{sec:appendixA}
%
Using $f(x) = \cos(x)$, we have
%
\[
f (x + i h) = \cos(x + i h),
\]
%
or through trigonometric identities equivalently
%
\[
f(x + i h) = \cos(x) \cosh(h) - i \sin(x) \sinh(h),
\]
%
Applying equation (\ref{eqn:complexFirstDeriv})
\[
\frac{\partial f(x)}{\partial x} = - \frac{\sin(x) \sinh(h)}{h},
\]
Since $h$ is arbitrary we desire it to be as small as possible or
\[
\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \sinh(h)}{h},
\]
using L'H\^opital's rule
\[
\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \cosh(h)}{1},
\]
of course, as $h \to 0$ then $\cosh(h) \to 1$ thereby recovering the exact derivative
\[
\frac{\partial f(x)}{\partial x} =-\sin(x).
\]

@@

\end{enumerate}
\label{doingthework}
%This shows how information flows through the nested functions in one expression, like a table of contents
For completeness we write out the computer's steps to evaluate \ref{differentiation} under the conditions of
\ref{abcvalues} with a particular value of $s = x_0$, in equation format: \\ \\
$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
%Repeating above in tabular format: \\ \\
%This shows the same information but is ordered to show progression, like chapters
%\begin{tabular}{l l l}
%Current Evaluation & $s$ Value & Partial Derivative \\ \hline
%$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
%$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
%$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
%\end{tabular} \\ \\
%This explains what the equation and table above mean.
Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand,
all it needed to do was:
\begin{enumerate}
\item to evaluate each of $u,v,w$ according to \ref{abcvalues}, in order from $w \rightarrow
v \rightarrow u$, 
\item to remember the values for $x_0$ and the output of each function evaluation besides  $u$, 
\item then to use $x_0$ and the output of the function evaluations as input for the corresponding partial
derivative function  evaluations, and
\item store the individual partials.
\end{enumerate}
Lastly, to compute the partial derivative of the entire composition function, the computer multiplies
the individual partials together in observance of the chain-rule.

>>>>

The methods were compared by running two sets of test problems where each
method would solve a problem concurrently. In this context, concurrently means
successively, drawing from the same independent variables, yet within the same
computational process; this definition is characterized in further detail in
Section~\ref{JGAM}. Each test problem simulated a $\SI{4}{\meter}$ by
$\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing
tension along the axis parallel to the long dimension of the block. The test
problems were set up in \emph{Perdigm} as quasistatic equilibrium problems,
where the displacement boundary conditions used to apply tension were applied
gradually in ``load steps'', and an equilibrium solution for each load step was
achieved before applying the next load step. The purpose of applying the load
in gradual steps is a widely used technique to ensure the Newton's method is
always initialized near the actual solution and therefore likely to converge.
The fictional material represented in the model had values of
$\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$
for shear modulus. The peridynamic horizon, a length scale that effectively
sets the bandwidth of tangent-stiffness in the computation, used in the test
problems was always set to three times the nominal discretization size (i.e.
node spacing in the particle discretization scheme). This results in a minimum
tangent-stiffness matrix bandwidth of 21 for a 3D problem (for an optimal node
numbering scheme). Pictured in Fig.~\ref{fig:SCMesh} is a viewgraph for one of
the peridynamic meshes used in this study. The coloration in the image
indicates node numbering, where a smooth color transition means that the nodes
have close ID numbers. The abrupt color changes in the center and at the ends
of the body show that numbering is not contiguous and is assigned with respect to four regioins. The purpose of designating
these regions, for this study, was to apply displacement boundary
conditions called \emph{volume constraints} in peridynamics.

@@

\begin{table*}[!ht]    
  \scriptsize
  \centering
        \caption{Test Parameters} \label{tab:TestParams}   
       \begin{tabular}{c c c c c c}
         \toprule
         Peri. nodes & Nozero TS elems. & Horizon & Load rate & # Load steps & Abs. tolerance \\ 
                     & $\times 10^6$    & \si{\meter}& \si{\centimeter\per\second} &   &     \\
        \midrule
        1000 & 1.99             & .3   & 2.0 & 2 & 10^-8\\
        2000 & 4.25             & .24  & 2.0 & 2 & 10^-8\\
        3000 & 7.79             & .21  & 2.0 & 2 & 10^-8\\
        4000 & 15.4             & .19  & 2.0 & 2 & 10^-8\\
        6000 & 20.0             & .16  & 2.0 & 2 & 10^-8\\
        8000 & 31.4             & .15  & 2.0 & 2 & 10^-8\\
        12000 & 40.1            & .131 & 2.0 & 2 & 10^-8\\
        16000 & 83.9            & .12  & 2.0 & 2 & 10^-8\\
        32000 & 165.0           & .095 & 2.0 & 2 & 10^-8\\
        1000000 & 1670          & .03  & 2.0 & 2 & 10^-8\\
        \bottomrule
    \end{tabular}
\end{table*}

@@

Data for accuracy measurements were taken according to the methods mentioned in
Section~\ref{tcs}, and reduced using Eqn.~(\ref{eqn:accuracy}). The
averaged results of these accuracy measurements for the serial tests can be
seen in Fig.~\ref{fig:serial_accuracy}. Averaged Frobenius norm (referred to as
$l^2$-norm in the figure labels for conciseness) of the element-wise difference
between a given TS matrix and the TS matrix produced by the AD method is
plotted as a function of the number of nonzero TS matrix elements. CS was shown
to be the most accurate of the methods when compared to AD and was roughly
\emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of
magnitude more accurate than FD}. This accuracy ranking order was maintained
for each test in the series. It should be pointed out that compared to the
$l^2$-norms of any of the TS matrices by themselves, the magnitude of the
accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods
is relatively small. Relative error was computed from these measurements and is reported in
Table~\ref{tab:RelAcc}. The values for the $l^2$-norms of the TS matrices appear
in the simulation data which is available in the paper repository. The effect
of these accuracy differences on convergence rate of the Newton solver was studied and the
results are reported in section~\ref{sec:PeridigmConvergenceStudy}.  
%
The accuracy of FD and CD could be improved by using a smaller probe distance,
$h$; however, if $h$ is too small, then the methods can suffer from severe
effects, including solution divergence, of the round-off error associated with
subtracting two numbers that are very close to one another.  The value of $h$
used in this study is the default value used in \emph{Peridigm} which is
heuristically derived based on computational node spacing using the method explained in \cite[pp. 90]{ref-Adaggio}.  
\add[Michael]{To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
as is not a user determined setting in \emph{Peridigm}.} One of the beauties
of the CS method is that since dependence on $h$ decreases for small $h$, which
a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$,
cannot be too small, therefore it can be set to machine epsilon and the
analysis performed without the issues associated with round-off error. The
value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to
demonstrate this.
%
\begin{figure}[tbp] \centering
\scalebox{1.0}{\input{./figs/serial_accuracy.pgf}} \caption{Serial test series
accuracy measurements.} \label{fig:serial_accuracy} \end{figure}
%
The averaged results of the accuracy measurements taken for the parallel test
series can be seen in Fig.~\ref{fig:multi_accuracy}. The ordinate axis units
remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the abscissa now
indicates number of compute cores used to solve the test problem. Accuracy
trends matched those for the single core tests. 
%
\begin{figure}[tbp] \centering
\scalebox{1.0}{\input{./figs/multi_accuracy.pgf}} \caption{Multicore test
series accuracy measurements.} \label{fig:multi_accuracy} \end{figure}
%
It can be seen in Figs.~\ref{fig:serial_accuracy} and ~\ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods. Those figures are depicting absolute error. If relative error is computed from the absolute error measurements by dividing them by the Frobenius norm of the appropriate AD produced TS matrix, it can be seen as in Table~\ref{tab:RelAcc}, that there is little variation in the accuracy of the methods between tests or between serial and parallel test groups.

\begin{table*}[!ht]    
  \scriptsize
  \centering
        \caption{Relative accuracy results} \label{tab:RelAcc}   
       \begin{tabular}{c c c c c}
         \toprule
         Cores & TS Elements & \multicolumn{3}{c}{Percent error} \\ 
         & $\times 10^6$ & CD & CS & FD \\
        \midrule
        1 & $1.99$  & 1.28x10^-8 & 2.29x10^-15 & 1.46x10^-5\\
        1 & $4.25$  & 8.86x10^-8 & 2.25x10^-14 & 1.34x10^-5\\
        1 & $7.79$  & 1.37x10^-8 & 2.27x10^-14 & 1.25x10^-5\\
        1 & $15.4$  & 4.33x10^-8 & 2.52x10^-14 & 1.16x10^-5\\
        1 & $20.0$  & 7.97x10^-8 & 2.18x10^-15 & 1.13x10^-5\\
        1 & $31.4$  & 1.44x10^-8 & 2.29x10^-15 & 1.09x10^-5\\
        1 & $40.1$  & 1.36x10^-7 & 2.24x10^-15 & 1.05x10^-5\\
        1 & $83.9$  & 1.30x10^-7 & 2.44x10^-14 & 9.63x10^-6\\
        1 & $165.0$  &1.30x10^-7 & 2.44x10^-14 & 9.63x10^-6\\
        32 & $1670$  &6.00x10^-7 & 2.29x10^-14 & 6.93x10^-6\\
        64 & $1670$  & 6.01x10^-7 & 2.31x10^-14 & 7.02x10^-6\\
        96 & $1670$  & 6.03x10^-7 & 2.33x10^-14 & 7.10x10^-6\\
        128 & $1670$  & 5.97x10^-7 & 2.34x10^-14 & 7.19x10^-6\\
        \bottomrule
    \end{tabular}
\end{table*}

@@

\section{References}

@@

\appendix
\renewcommand*{\thesection}{\Alph{section}}
%
\section{Complex-Step Example}
\label{sec:appendixA}
%
Using $f(x) = \cos(x)$, we have
%
\[
f (x + i h) = \cos(x + i h),
\]
%
or through trigonometric identities equivalently
%
\[
f(x + i h) = \cos(x) \cosh(h) - i \sin(x) \sinh(h),
\]
%
Applying equation (\ref{eqn:complexFirstDeriv})
\[
    \frac{\partial f(x)}{\partial x} = - \frac{\sin(x) \sinh(h)}{h} + \mathcal{O}(h^2).
\]
Since $h$ is arbitrary we desire it to be as small as possible or
\[
\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \sinh(h)}{h},
\]
using L'H\^opital's rule
\[
\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \cosh(h)}{1},
\]
of course, as $h \to 0$ then $\cosh(h) \to 1$ thereby recovering the exact derivative
\[
\frac{\partial f(x)}{\partial x} =-\sin(x).
\]

@@

\end{enumerate}
\label{doingthework}
%This shows how information flows through the nested functions in one expression, like a table of contents
For completeness we write out the computer's steps to evaluate \ref{differentiation} under the conditions of
\ref{abcvalues} with a particular value of $s = x_0$, in equation format: \\ \\
$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
%Repeating above in tabular format: \\ \\
%This shows the same information but is ordered to show progression, like chapters
%\begin{tabular}{l l l}
%Current Evaluation & $s$ Value & Partial Derivative \\ \hline
%$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
%$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
%$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
%\end{tabular} \\ \\
%This explains what the equation and table above mean.
Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand, which of course needed to be specified using an AD data structure from a library like Trilinos::Sacado \cite{trilinos} or ADIFOR \cite{bischof1995adifor},
all it needed to do (for 'backward' AD) was:
\begin{enumerate}
\item to evaluate each of $u,v,w$ according to \ref{abcvalues}, in order from $w \rightarrow
v \rightarrow u$, 
\item to remember the values for $x_0$ and the output of each function evaluation besides  $u$, 
\item then to use $x_0$ and the output of the function evaluations as input for the corresponding partial
derivative function  evaluations, and
\item store the individual partials.
\end{enumerate}
Lastly, to compute the partial derivative of the entire composition function, the computer multiplies
the individual partials together in observance of the chain-rule.
